---
title: "index"
authors: "Steven Tran, Eric Large, Jinsu Ha, Jason Lim"
date: "2/18/2022"
output: html_document
---
# Code Name: BBB

# Winter 2022

# Affiliation:
**Info 201: Technical Foundations of Informatics (The Information School)**
**THE UNIVERISITY OF WASHINGTON**

# Project Title

NBA Analysis

# Audience

Our main demographic includes future employers, basketball fans, and perhaps most importantly, basketball players themselves and the owners of these NBA teams. 

# Abstract

Our data analysis project mainly focuses on the success and health of current 
and former NBA players, with emphasis on the 2021 season as well as the 
question, "What is the impact of a player's health on their performance?". 
To accomplish this we will analyze a multitude of player statistics, including 
but not limited to, All-Star Game appearances (to operationalize success), years 
played, and injuries, comparing seasonal and career longevity with overall 
player success. Finally, conducting this insightful analysis is important 
because it will help shed light on various real-world issues including the 
mental and physical health of athletes, along with modern day expectations on 
individual performance in professional basketball.

# Key Words
<li>NBA stats summary</li>
<li>Player’s performance</li>
<li>Career Longevity</li>
<li>Health</li>
<li>Points</li>

# Introduction

It seems as though every action, every play, and every call on the basketball 
court is compiled and recorded as a measurement of the essence of a player. It’s
what defines them. Fans might use it to back up their argument that Aleksej 
Pokuševski is the greatest player alive or NBA teams may use it to gauge the 
value of players. For our group, we wanted to investigate the relationship 
between the performance of the players and their physical and mental health. 
These two metrics deeply impact the players, whose income depends on their 
ability to perform, the owners, as the players create the bulk of their revenue,
and finally the NBA fans that look forward to seeing their favorite players lace
up. Some possible ways we can approach this would be to analyze CSV files on 
each player’s minutes played, fouls drawn, points per game, and efficiency of 
each shot to determine if there is a relationship. For instance, statistics that
indicate health and longevity such as games played and minutes per game would be
analyzed with their production of points and how efficiently the players are 
getting those points. These findings may impact real world athletes and how they
can improve their training regimen and our goal is to identify risks that may 
pop up from our analysis.

# Design Situation

Our main topic of concern is that we want to know the relationship between the 
success and the health of current and former NBA players in 2021. Key elements 
of the topic include: taking care of their mental and physical health, and how 
these relate to their performance. Some questions that we can ask are: "How do 
NBA players continue to strive for success and not get distracted with always 
having so much pressure and attention on them?" & "How does the success of a 
player and their career longevity correlate?"

Human values that are connected to the topic of our concern are personal 
development, mental health, and strength. As we are focusing on each individual 
player’s health and longevity that lead to successful/unsuccessful performance 
in the NBA, these values are significant because they have direct correlations 
to each other.

The stakeholders of our dataset would be first and foremost the players whom 
we collect data on, the owners of the NBA teams, and any basketball fan in 
general. The health and production of these players are fundamentally important 
to all three of these groups as the NBA is a business at the end of the day. 
Indirect stakeholders may be athletes of other sports. If there is a correlation
with a metric of health and production, it could translate to other sports with 
similar intensities.

The potential benefits of this analysis include increased awareness around 
(potentially) unrealistic player performance standards and increasingly long 
careers through load management. By highlighting both injury and subpar 
performance patterns in players, our analysis supports the health and well-being
of professional basketball players. Finally, NBA players must constantly adapt 
to their aging bodies and dynamic environments associated with the game, which 
is an aspect that our project seeks to address for their aid.

Unanticipated consequences or harms of our analysis might include a deficiency 
in overall player statistics, as we are narrowing our focus on the 2021 season 
along with data compiled from the last ten years. Although it’s seemingly 
unethical to “predict” player injuries, we find it necessary to analyze their 
minutes, games played, and other stats leading up to previous injuries to 
prevent future instances. Beyond these factors, we will attempt to be as 
unbiased and inclusive as possible in our analysis.

# Research Questions

What effect(s) does a conservative approach (e.g., players with high game score 
playing relatively low minutes; see Kawhi Leonard) have on a player’s overall 
efficiency for that year? By comparing a players game score (GmSc), minutes 
played (MP), and field goal percentage (FG%) stats, we can contextualize a star 
player’s efficiency given their work load.

How does average number of fouls affect a player’s performance during the 
season?

We could access the average number of fouls a player has received in a single 
NBA season and plot that with their true shooting percentages, like in Nikhil
Ajjarapu's data set or compare these values with the total points per game in 
that season. This is a topic of interest since fouls are the main indicator of 
how rough or strenuous it is in a basketball game. It is assumed that the health
of a player that receives on average 3 fouls a game is in more jeopardy than a 
player who receives 1.

How does the frequency of injury relate to the game performance?

Analyzing games played in a season with their efficiency would be a good 
indicator how their games missed due to injury impacts their play.

What effect does it have on bench players’ performance when they start the game
instead of the original starters who are healthy?

By comparing bench and starters, we can better understand how much the increased
time spent on court impacts their health.

What players have the largest decrease in performance measured by points per 
game, true shooting percentage, and minutes played?

As said in the question, we can analyze the metrics of performance such as 
points per game, true shooting, with metrics of health such as minutes played 
or time spent recovering from injury. This is perhaps the crux of our research 
and something we'd like to draw conclusions from.


# Expected Implications
Our expected implication is that the healthier the players are, meaning less 
injuries, more years of playing, and more overall successful seasons, the more
successful the player can be.

This would ultimately be best for all parties involved as fans get to see their
favorite players play for longer, the athletes get more opportunities to earn a
salary doing what they love, and NBA owners get to see quality players that 
attract more fans to their arenas.

If our research question gets answered and we get correlations between minutes 
played and production, the implications for the NBA regular season are massive.
Perhaps it is time to rethink the number of games to preserve the quality of 
the game.

# Limitations
Though our analysis will consist of the appropriate amount of data points to 
best support our thesis and come up with results, the variation of players’ 
performance is expected as their progress is not always linear. For example,
players go through injuries, personal matters, or slumps, and later some of 
those players show good performance in the game but there are also others who
struggle to be at the same level as they were before. Thus, we’d like to 
emphasize that the end result of the project will be based on the probability 
from the collected data and won’t be entirely accurate in future cases. 
Additionally, if our group decides to take players’ injuries into consideration 
for the project, we’ll need to assess the severity of each player’s injury as 
well as they matter to their future performances, which would be difficult.
# Acknowledgements

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r summrary information, echo = TRUE, message = TRUE}
source("Summary_Information.R")


```

```{r table of summary information, echo = TRUE}
source("Table_Summary_Information.R")
kable()
```

```{r chart 1, include = TRUE}
source("Chart_1.R")
plot(chart_1)
```
# Purpose of Chart 1:
The purpose of this dot plot is to demonstrate the relationship between
the number of games and points scored by each team in the NBA. As it's 
hard to show all data points, which is approximately over 750 data points, the 
new data frame is grouped by team which makes it easier to plot. Also,
by grouping the players by team, we are able to see the differences among these 
teams. 

# Observation & Insights:
From this visualization, we have come to a conclusion that the more games the
players played, less average points were recorded. Looking at the maximum
of the smooth line from the plot, the highest points were observed around 
55 games and 68 games. However, as the number of average games increase, there was
somewhat rapid decrease towards the total number of games in one season. 

```{r chart 2, include = TRUE}
source("Chart_2.R")
plot(chart2)
```
# Purpose of Chart 2:
The purpose of including this bar chart in our report
is to highlight the modern 3-point efficiency (3P%) of each 
position type in the NBA. Although our data set includes
only active starters for all 30 teams during the 2021-22 season,
it still provides a large sample of current 3-point efficiency 
ratings at each of the major five positions. 

# Observations & Insights:
From this visualization, we capture a snapshot on the current standard
of 3-point shooting efficiency as it relates to every position. As we have
seen since the turn of the century, the modern NBA has adapted to the long-range
shooting style-of-play more than ever, reflected in these averages. Despite not 
having insight into the center (C) position (due to a lack of data from the 
2021-22 season), we can still infer that the wing positions (SF, PF) having 
relatively similar averages (33.1% and 33.3%) to the point guard (PG) position
(35.1%) means there has been an emphasis placed on this aspect of the game for
every player, regardless of position. This shows the high level of skill each
individual must have to even be considered "average" for their position in the
modern NBA. When considering the context of the 3-point shot,it is hard to 
imagine a time where the 3-point line did not exist.

```{r chart 3, include = TRUE}
source("Chart_3.R")
plot(chart_3)
```
# Purpose of Chart 3:
The purpose of Chart 3 is to analyze the most injured team this season and to 
identify the correlation between the average minutes played by each player on 
the Orlando Magic Vs. the average effective field goal percentage for the team
that season. This would help us understand the effect of health and injuries 
on the effectiveness of NBA players. Since the Orlando Magic was the most 
injured team, it made sense to understand how those injuries impacted their 
play.  
We chose a line graph because the position of the dots and the direction
of the lines make it easier to understand this correlation. 

# Observations & Insights:
It seems as though the team is most effective when players average from 27 to 
30 minutes per game. The effective field goal percentages start to plateau and 
then dip to a low of 0.425 when players average about 40 minutes. 
This is staggering because the more players get injured, the less effective 
they are on court, and this may lead to even more injuries. 

setwd("~/Documents/Info201Assignments/final-project-starter-OrangeCowz/source")


